CS6999 Web Data and Text Mining:\\Parallel Text Indexer Project Report

### 2011-05-16 report/report.tex
First draft of literature survey complete

%----------------------------------------
% Introduction
%----------------------------------------
\section{Introduction}
Fundamental to the real-world application of \textit{Information
  Retrieval} techniques is the ability to objectively measure the
quality of a particular algorithm or system and compare it to
others. So often, information retrieval algorithms lack an analytical
model that would allow one to make formal statements as to the
performance boundaries of a particular system. In addition, as with
machine learning and classification problems, using a scalar value to
represent the performance of a system as a whole is misleading and
does not account for the variety of operating conditions under which
the system may be expected to perform.

@@

    \item The storage requirements of the technique. How do the
      storage requirements evolve with the size of the collection
      and/or degree of concurrency employed ?
\end{itemize}

To drive information retrieval research in one direction or another,
it would be useful to have at ones disposal a physical implementation
framework within which one could empirically explore the operating
parameters of a particular algorithm and/or technique along the lines
described above. Such a system would allow researchers to investigate
how the system responds to variation in any one of these operating
parameters. Does the query speed increase as more computational
capacity is added ? How do storage requirements vary with the size of
the document collection ? The aim of this project is to create such an
information retrieval \textit{test-bed}.

The initial reference implementation of this information retrieval
test-bed is based on a simple \textit{inverted index} but has been
architected such that the specific algorithm could easily be changed
without major rework. This paper presents a survey of existing
literature in support of our test-bed concept, as well as the specific
algorithmic techniques embodied in the reference implementation. In
addition, a detailed architectural overview of the system is
presented, giving insight into how the system would be modified for
alternative algorithms. Lastly, an evaluation of our inverted index
reference implementation is conducted and the results analyzed.

@@

The Aho-Corasick string matching algorithm\cite{RefWorks:103} is a
kind of dictionary matching search algorithm that constructs a finite
state machine to scan for a given set of keywords. It is, in effect, a
reduced grammar regular expression parser described in
\cite{RefWorks:111}.

>>>>

%----------------------------------------
% Introduction
%----------------------------------------
\section{Introduction}
Fundamental to the real-world application of \textit{Information
  Retrieval} techniques is the ability to objectively measure the
quality of a particular algorithm or system and compare it to
others. Often, information retrieval algorithms lack an analytical
model that would allow one to make formal statements as to the
performance characteristics of a particular system. In addition, as with
machine learning and classification problems, using
a scalar value to represent the performance of a system as a whole is
misleading and does not account for the variety of operating
conditions under which the system may be expected to perform.

@@

    \item The storage requirements of the technique. How do the
      storage requirements evolve with the size of the collection
      and/or degree of concurrency ?
\end{itemize}

To drive information retrieval research in one direction or another,
it would be useful to have a physical implementation
framework within which one could empirically explore the operating
characteristics of a particular algorithm and/or technique along the lines
described above. Such a system would allow researchers to investigate
how the system responds to variation in any one of these operating
parameters. The aim of this project is to create such an
information retrieval \textit{test-bed}.

The initial reference implementation of this information retrieval
test-bed is based on a simple \textit{inverted index} but has been
architected such that the specific algorithm could easily be changed
without significant rework. This paper presents a survey of existing
literature in support of our test-bed concept, as well as the specific
algorithms employed in the reference implementation. In
addition, a detailed architectural overview of the system is
presented, giving insight into how the system would be modified for
alternative algorithms. Lastly, an evaluation of our inverted index
reference implementation is conducted and the results analyzed.

@@

\subsection{Indexing}
\textit{Information Retrieval} is a fast becoming the dominant form of
information access, overtaking traditional relational database style
searching \cite{RefWorks:109}. The basis of most document retrieval
systems is the \textit{term-document} incidence matrix. This matrix is
typically sparse and more efficiently represented as an
\textit{inverted-index} which maps terms to the parts of a document
where they occur.

In order to construct an \textit{inverted index}, documents must be
scanned to determine \termit{term frequencies}. The Aho-Corasick
string matching algorithm\cite{RefWorks:103} is a simple and efficient
text scanning algorithm. The algorithm constructs a finite
state machine to scan for a given set of keywords. It is, in effect, a
reduced grammar regular expression parser of the type described in
\cite{RefWorks:111}. The algorithm is simple and efficient,
construction time being proportional to the sum of the length of the
keyword set, and the number of state transitions required to scan a
document is independent of the number of the size of the keyword
set.

\subsection{Evaluation}
In the case of machine learning and classification systems, the
specification of a particular set of misclassification costs or false
and positive misclassification rates define a particular
\textit{operating point} of the system.

This concept applies equally to \textit{information retrieval} systems
whereby it is misleading to represent the operating characteristics of
a particular system with a scalar metric. In addition, the evaluation
of information retrieval systems suffers from the the inherent
subjectivity with respect to \textit{relevance}
\cite{RefWorks:72}. Non-scalar evaluation methods try to characterize
the performance of a system over a range of operating
conditions. Although this makes direct performance comparisons
ambiguous, we are more likely to be able to select the algorithm which
will perform best in the real world by applying domain expertise to
the results of a non-scalar evaluation \cite{RefWorks:32}.

\subsection{Parallelization}
Once an objective measure has been established, the performance of an
information retrieval system can be improved iteratively by making
various optimizations. These optimizations may take the form of
algorithm enhancements, implementation efficiencies, or
parallelization.

Algorithm enhancements usually represent the forefront of research in
the field. Work to validate a new concept or algorithm is usually
performed in a very controlled environment and on relatively small
document collections. As the field matures, improvements are harder
fought and often involve increasingly complex algorithms and corpus
approximation methods into which it becomes more difficult to
\textit{fold-in} new documents. In addressing the considerable
computation cost of re-calculating \textit{LSI} \cite{RefWorks:102}
approximations, \cite{RefWorks:101} introduced new update methods
which saw improved average precisions but were forced to conclude that
the improvements were at a computational premium over previous
methods. A semi-discrete matrix decomposition for LSI, \textit{SDD},
performs as well as the original SVD-based LSI method for equal query
times and uses less storage \cite{RefWorks:99}. The decomposition time
of the mew method, however, is significantly larger than the original.

Enhancement of implementation efficiency in \cite{RefWorks:96} saw a
six-fold timing improvement in \textit{LSI} searches on large document
collections. The same study achieved nearly 180 times improvement
through the use of parallelization. The study made no algorithmic
improvements and the result is consistent with increased use of
parallelization by large commercial systems to cope with the ever
increasing bodies of data that must be processed.

\textit{Map-Reduce}, popularized by Google, is emerging as an
important programming model for large-scale data-parallel applications
such as web indexing \cite{RefWorks:113}. Fundamentally, Map-Reduce
breaks up a computation into smaller tasks which are made to run in
parallel on different physical machines. The \textit{map} and
\textit{reduce} phases are inspired by the functions of the same name
used in \textit{functional programming}. The model scales well to large
clusters mostly because of the decoupled nature of the tasks
themselves. An initial \textit{map} phase breaks the data into
  separate chunks and assigns them to a task. The \textit{reduce}
  phase assembles the results of the map tasks to produce the final
  result.

\textit{Hadoop} is an open source implementation of the MapReduce
model, developed mostly by Yahoo, and now maintained by the Apache
Software Foundation. Hadoop can be configured to automatically create
speculative copies of slow running map tasks
(stragglers). This facility is intended as a pre-emptive failure
recovery mechanism as well as a means of optimizing overall
throughput. The metrics used to identify stragglers as well as the
algorithm for determining the target node on which speculative copies
of a task should be launched, is examined in \cite{RefWorks:128}. The
paper introduces an alternative scheduling algorithm, \textit{LATE},
to address excessive task speculation issues with the native Hadoop
scheduler in a heterogeneous hardware environment. Experiments
comparing the native Hadoop scheduler, a non-speculative scheduling
strategy, and LATE were conducted using hosted virtual machines (VMs)
in the Amazon EC2 environment as well as a dedicated local
cluster. Variance in node performance was introduced by varying the
distribution of VMs across physical machines and also introducing
network and disk transfer loads. Results show speculative execution
(native Hadoop and LATE) outperforming non-speculative execution and
LATE outperforming native Hadoop in most cases.
